Implicit yaddayadda

### 2011-10-28 papers/PLDI2012/implicitization.tex
llvm + apron

Most SMT solvers, including Z3,%
\footnote{\url{http://research.microsoft.com/en-us/um/redmond/projects/z3/}}
Yices,%
\footnote{\url{http://yices.csl.sri.com/}}
and all those supporting the full SMTLIB2 standard \cite{BarST-SMTLIB},
offer an \emph{incremental} interface: the client program specifies the formula as an initially empty conjunction, to which additional constraints are added, and calls a ``check'' function answering whether it is satisfiable; 
it may then backtrack some of the constraints and add other ones without restarting from scratch.

@@

The $\sigma$ function takes as parameters the index of the starting abstract
value, and the path we focus on. If we dynamically construct this function
during the analysis, we should store the results of the $\sigma$ function into a
compact representation, such as Algebraic Decision Diagrams.
\DM{Pourquoi des ADD et pas des BDD?}
Note that $\sigma(j,k)$ will be computed only if needed.

@@

We have implemented our proposed solutions inside a prototype intraprocedural
static analyzer,
as well as the classical abstract interpretation algorithm, and the state-of-the-art
techniques \emph{Path Focusing} \cite{Monniaux_Gonnord_SAS11} and \emph{Guided
Static Analysis} \cite{DBLP:conf/sas/GopanR07}.
We also can choose among different abstract domains: convex polyhedra, octagons,
intervals, grids (convex polyhedra and linear congruences).
For SMT-solving, our analyzer uses Yices
\cite{DBLP:conf/cav/DutertreM06} or Microsoft Z3\cite{DBLP:conf/tacas/MouraB08}
as SMT-solver.

@@

\subsection{Analysis algorithm}
\label{sec:analysis-algorithm}
Our tool operates over LLVM bitcode \citep{LLVM_langref,Lattner:2004:LCF:977395.977673}. LLVM bitcode is in \emph{single static assignment} (SSA) form: a given scalar variable is given a value at a single syntactic point in the program. In concrete terms, an assignment \lstinline|x=2*x+1;| gets translated into a definition $x_2 = 2x_1+1$, with distinct variables $x_1$ and $x_2$ corresponding to the same original variable \lstinline|x| at different points in the program.
LLVM makes it easy to follow definition-use and use-definition chains: for a given variable (say, $x_2$) one can immediately obtain its definition (say, $2x_1+1$).
One may see conversion to SSA form as a static precomputation of some of the symbolic propagations proposed by \citet{DBLP:conf/vmcai/Mine06} to enhance the precision of analyses.

>>>>

Most SMT solvers, including Z3,
%\footnote{\url{http://research.microsoft.com/en-us/um/redmond/projects/z3/}}
Yices,%
%\footnote{\url{http://yices.csl.sri.com/}}
and all those supporting the full SMTLIB2 standard \cite{BarST-SMTLIB},
offer an \emph{incremental} interface: the client program specifies the formula as an initially empty conjunction, to which additional constraints are added, and calls a ``check'' function answering whether it is satisfiable; 
it may then backtrack some of the constraints and add other ones without restarting from scratch.

@@

The $\sigma$ function takes as parameters the index of the starting abstract
value, and the path we focus on. If we dynamically construct this function
during the analysis, we should store the results of the $\sigma$ function into a
compact representation, such as Algebraic Decision Diagrams.
Note that $\sigma(j,k)$ will be computed only if needed.

@@

We have implemented our proposed solutions inside a prototype intraprocedural
static analyzer,
as well as the classical abstract interpretation algorithm, and the state-of-the-art
techniques \emph{Path Focusing} \cite{Monniaux_Gonnord_SAS11} and \emph{Guided
Static Analysis} \cite{DBLP:conf/sas/GopanR07}.

Our tool operates over LLVM bitcode \citep{LLVM_langref,Lattner:2004:LCF:977395.977673}, which is a target for several compilers, most notably Clang (supporting C and C++) and llvm-gcc (supporting C, C++, Fortran and Ada).
Abstract domains are provided by the APRON library \citep{DBLP:conf/cav/JeannetM09}, and include convex polyhedra, octagons, intervals, and linear congruences.
For SMT-solving, our analyzer uses Yices
\cite{DBLP:conf/cav/DutertreM06} or Microsoft Z3\cite{DBLP:conf/tacas/MouraB08}.

@@

\subsection{Analysis algorithm}
\label{sec:analysis-algorithm}
LLVM bitcode is in \emph{single static assignment} (SSA) form: a given scalar variable is given a value at a single syntactic point in the program. In concrete terms, an assignment \lstinline|x=2*x+1;| gets translated into a definition $x_2 = 2x_1+1$, with distinct variables $x_1$ and $x_2$ corresponding to the same original variable \lstinline|x| at different points in the program.
LLVM makes it easy to follow definition-use and use-definition chains: for a given variable (say, $x_2$) one can immediately obtain its definition (say, $2x_1+1$).
One may see conversion to SSA form as a static precomputation of some of the symbolic propagations proposed by \citet{DBLP:conf/vmcai/Mine06} to enhance the precision of analyses.
